### 2020-05-28 referee-report-20200409-response.txt
Here is the version going out to the co-authors for pre-approval before getting returned to the referee

Dear editor & referee,

We would like to thank both of you for encouraging us to look more deeply into the justification of the synthesis filter that we were using. The result has been not only an improved paper (hopefully!) but also an improved system. We found it interesting to compare the performance of different kinds of filters, and hope that the extra rigour we've inserted into the "Synthesis filters" section will be acceptable to the referee and useful to the readers. On account of the extra work, several sections of the paper have been completely rewritten, as described below in the detailed comments.

----------------------------

Referee: This paper present a method for inversion of the 2nd-stage (fine) PFB on data recorded with the Murchison Widefield Array's Voltage Capture System, in order to recover high time resolution instead of the VCS' nominal fine (10kHz) frequency resolution. The authors demonstrate the performance on a trio of pulsars with known short-timescale emission structure. I appreciate very much the effort the authors went to to write up this work, and generally feel that more papers of this sort, covering signal analyses suitable suitable to radio observations, are needed in the literature. Unfortunately, as it stands, I don't see the novelty or scientific impact of this work, and cannot recommend it for publication in a peer-reviewed science journal. I strongly encourage the authors to flesh out their derivations, to justify and characterize the implementation more quantitatively, and re-submit. Detailed comments follow.

Referee: The paper begins with a concise review of PFBs. An overview of synthesis filters used to reverse their operation is given, and the problem is overall well motived. I applaud the authors for laying out the task so clearly.

Referee: The specific choice of synthesis filter, however, seems completely unmotivated and unjustified in any quantitative sense. The authors seem to simply argue "good enough is good enough," but for the central point of the paper I find this woefully insufficient, particularly when well-justified formulations are easily found online and in the community. A quick google reveals https://github.com/jrs65/pfb-inverse as the top result, where a least-squares solution to the inversion is presented. It appears to operate offline, but solves the artefacts described in this work. If the authors require realtime inversion to make it a part of the MWA VCS pipeline, computationally inexpensive solutions exist which vastly reduce the temporal echoes they see. Given that such solutions are in the community already, I would expect at minimum some proof of optimality for some well-chosen metric.

Authors: In hindsight, we completely agree that a paper such as this, if it does not attempt to provide a proper justification for the choice of synthesis filter, cannot possibly offer any value to the radio astronomy community who might be interested in replicating, or improving upon, our method. Accordingly, we investigated the least-squares solution that the referee recommended, and have provided a full, quantitative comparison of multiple synthesis filters, including our original "mirror" filter and a few least-squares solutions of different lengths. In so doing, we inevitably found that the least-squares solution performs better than our original filter, and we provide a metric (with accompanying derivation) for comparing them. Armed with this analysis, we implemented (one of) the least squares filters in our software, which is now available. All of this is reflected in the completely written Sections 3 and 4.

Authors: The analysis also revealed that our earlier analysis of the temporal echoes was flawed. Our previous estimates of the magnitude of this effect were greatly exaggerated, and all subsequent test show that the temporal echoes that we see are consistent with the $\lesssim -25$ dB "ringing" shown in the bottom panel of (the new) Fig. 4, which is a general feature of all imperfect back-to-back systems. Accordingly, we no longer had any reason to suppose that any of the features we saw in our pulsar observations were due to this effect, and have removed the relevant sentences from that section.

Referee: I was glad to see the impulse response to characterize performance, but the comparison to an ifft seems silly except to point out that one should not expect iffts to invert pfbs.

Authors: Yes, we agree. The comparison with the ifft is not useful, and we have removed those paragraphs. However, the referee will note that the new version no longer contains an impulse response test. The justification for this decision we give in the paper itself (the second paragraph of Section 4), which amounts to an argument that the impulse response is not the best way to characterise the performance of our particular system (whose implementation introduces significant non-linearities). We therefore focus our efforts on the empirical test done on actual data (Section 4), being careful to quantify its performance in terms of the same metric introduced earlier (Section 3.1).

Referee: The pulsar observations constitute a nice proof-of-concept, and serve as demonstrations of the chosen filter in the field as the authors do not pursue them to novel results. I do not understand the comparison of incoherently- to coherently-dedispersed profiles, it seems very much apples-to-oranges and not directly relevant to the variable time/freq division of the data. The time-resolved sub-pulse microstructure of B0950+08, however, is an impressive demonstration, well worth including in a revised paper.

Authors: The relevance of the comparison between coherent and incoherent de-dispersion lies in the fact having a high time resolution is useless if the resolution is limited by dispersive smearing across a fine channel, which can easily become the dominant factor for MSPs de-dispersed incoherently. In the original (non-high-time resolution) mode, coherent de-dispersion was not an option, and we wanted to showcase the difference. However, the point is well taken: this has more to do with the limitations of the downstream analysis, and not so much with the synthesis system itself. In order not to belabour that point, we have removed the incoherently de-dispersed version of the J2241−5236 plot (Fig 7); however, we have left both J0437-4715 plots in the paper (Fig 8), in order to help the reader identify the "notches" discussed in the text.

Referee: Lastly, I would be remiss if I didn't express some concern that the PFB implemented on the FPGAs cannot be exactly reproduced on general-purpose compute hardware! The difference in hardware can hardly matter for a digital operation, and myriad libraries exist to implement arbitrary bit-depth arithmetic in CPUs. It's clear from the properties presented in appendix A that a number of awkward features are already reproduced, and I'd strongly encourage the authors to find an FFT implementation (or implement one themselves) which perfectly mimics that on the FPGA.

Authors: Finding and running the correct version of the fixed-point FFT implemented on the FPGAs proved difficult: that version is no longer supported, and our early attempts to replicate it were not successful... although in honesty we cannot be 100% sure of the reason. Neverthess, this point may now be moot. If the referee accepts our arguments that the impulse response test does not adequately characterise the system, then there is no longer any need to reproduce the forward PFB. Instead, the empirical test we offer instead uses the actual implemented system, meaning that the reported performance includes any peculiarites introduced by the FPGA (related to the FFT or otherwise). Obviously it's less than satisfactory not to have a complete account of what the forward PFB is doing, but our investigations indicate that the rounding errors described in Section 4 and the Appendix account for virtually the entire error budget (beyond that introduced by the choice of synthesis filter itself). Any deeper investigation would only be warranted if there was some indication that such effects could be mitigated, which we don't find here.

Referee: I'll close by saying again that I think the literature needs more papers covering topics like this, but they need to be addressed in at least the depth available via my quick google search. Please motivate the design of your synthesis filter, show that it is in some sense the correct filter to choose: optimal for some metric, or with artefacts below the noise floor for plausible observations, or at least theoretically motivated somehow. I would be delighted to have a solid PFB inversion process in the formal literature, and would like to see the authors resubmit after a more careful exploration or justification of the filter choice.

Authors: We thank the referee again for this encouragement towards rigour, and hope that this latest version meets those standards!